05_DESeq_manipulation

### 2019-05-19 analysis/05_DESeq2_manipulate.Rmd
now with many loops and functions

## Make a PCA plot, heat map, and box plots for each tissue for each sex

```{r althings, eval = T}
## subset for test 
#m.colData <- m.colData %>%
#  filter(sextissue == "male_hypothalamus") %>%
# droplevels()

@@

  print(eachgroup)

  colData <- m.colData %>%
      dplyr::filter(sextissue == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- m.countData %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ treatment )

  print(dds)
  nrow(colData) > 50

  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
  print(dds)

  dds <- DESeq(dds) # Differential expression analysis

  vsd <- vst(dds, blind=FALSE) # variance stabilized

@@

# comapre all contrasts, save to datafrmes
totalDEGS=data.frame()
for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = "") #assigns usique rownames
      print(k)
      totalDEGS[k,1]<-i               
      totalDEGS[k,2]<-j
      totalDEGS[k,3]<- numDEGs(i,j) #caluculates number of DEGs
    }
  }
}

@@

allcontrasts <- totalDEGS %>%
  ggplot( aes(V1, V2)) +
    geom_tile(aes(fill = V3)) +
    scale_fill_viridis(na.value="#440154", 
                     limits = c(0, 3000),
                     breaks = c(0, 1000, 2000, 3000)) + 
    xlab(" ") + ylab("Timepoint") +
    labs(fill = "# of DEGs",
         subtitle = eachgroup)
plot(allcontrasts)

@@

for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = " vs ")
      print(k)
      results <- resvals2(i,j)
      DEGs <- cbind(DEGs,results)
    }
  }
}

>>>>

## Run DESeq on all subsets of the data

```{r dds}
## subset for test 
#m.colData <- m.colData %>%
#  filter(sextissue == "male_hypothalamus") %>%
# droplevels()

levels(m.colData$sextissue)

@@

  print(eachgroup)
  dds <- subsetDESeq(eachgroup)

  dfname <- paste("dds", eachgroup, sep = "")
  assign(paste("dds", eachgroup, sep = ""), dds)

  return(dfname)
}

```

## Create variance stabilized data

```{r vsd}

mydds <- list(ddsmale_hypothalamus) # for testing only
#mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
#              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)

for (eachdds in mydds){

  vsd <- vst(eachdds, blind=FALSE) # variance stabilized

  vdsname <- paste("vsd", eachgroup, sep = "")
  assign(paste("vsd", eachgroup, sep = ""), vsd)
  return(vdsname)
}

```

## Calculate and plot total DEGs

```{r totalDEGs}

@@

for (eachdds in mydds){

  # comapre all contrasts, save to datafrmes
  totalDEGS=data.frame()
  for (i in a){
    for (j in b){
      if (i != j) {
        k <- paste(i,j, sep = ".") #assigns usique rownames
        print(k)
        totalDEGS[k,1]<-i               
        totalDEGS[k,2]<-j
        totalDEGS[k,3]<- numDEGs(i,j) #caluculates number of DEGs
      }
    }
        b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
  }

@@

  allcontrasts <- totalDEGS %>%
    ggplot( aes(V1, V2)) +
      geom_tile(aes(fill = V3)) +
      scale_fill_viridis(na.value="#440154", 
                      limits = c(0, 3000),
                      breaks = c(0, 1000, 2000, 3000)) + 
      xlab(" ") + ylab("Timepoint") +
      labs(fill = "# of DEGs",
          subtitle = "eachgroup")
  plot(allcontrasts)
}
```

@@

for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = " vs ")
      print(k)
      results <- returnpadj(i,j)
      DEGs <- cbind(DEGs,results)
    }
  }
}
